---
title: "Anomaly Detection Analysis"
author: "Ann Mberi"
date: "7/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading some libraries


```{r}
library(dplyr)
library(magrittr)
library(knitr)
library(tidyverse)
```


**ANOMALY DETECTION**

# Reading and previewing the dataset
```{r}
ano <- read.csv("http://bit.ly/CarreFourSalesDataset")
head(ano)
```

This is to detect whether there are any anomalies in the given sales dataset.

# Checking the dimensions of our dataset
```{r}
dim(ano)
```

# Checking the structure of our dataset
```{r}
str(ano)
```


# Checking the Statistical summary
```{r}
summary(ano)
```

# Checking for null values
```{r}
colSums(is.na(ano))
```

# Changing the date column from character to Date
```{r}
ano <- transform(ano, Date = format(as.Date(Date, "%m/%d/%Y"), "%Y/%m/%d"))
ano <- transform(ano, Date = as.Date(Date))
sapply(ano, class)
```

# Grouping the dataset by the Date column
```{r}
Sales <- ano$Sales
Date <- ano$Date
ano = ano %>% arrange(Date)
```
 
## Lets preview our dataset 
```{r}
head(ano)
```

##Sales Distribution
### sort dates in ascending order
```{r}
ano_sales <- ano[order(ano$Date),]
head(ano_sales)
```

## Plotting Histogram to show sales distribution

```{r}
hist(ano$Sales)
```

# Transactions Count
# We’ll group the transactions per day then tally them
```{r}
ano_count <- ano %>% group_by(Date) %>% tally()
colnames(ano_count) <- c("Date", "Count")
head(ano_count)
```

## Visualizing Transaction count .



```{r}
library(anomalize)
```

```{r}
ano_count %>%
time_decompose(Count) %>%
anomalize(remainder) %>%
plot_anomaly_decomposition() +
ggtitle("Transaction Count: Anomaly Decomposition")
```

# Visualizing Transaction count.
```{r}
ano_count %>%
time_decompose(Count) %>%
anomalize(remainder) %>%
time_recompose() %>%
plot_anomalies(time_recomposed = TRUE, ncol = 3, alpha_dots = 0.5) +
ggtitle("Transaction Count: Anomaly Recomposition")
```