---
title: "IP - Part3 Association Rules"
author: "Ann Mberi"
date: "7/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



**ASSOCIATION RULES**

# Installing and reading the necessary packages for the association rules analysis
#install.packages("arules", dependencies = TRUE)


```{r}
library(arules)
```

# Reading and previewing the dataset as transactions
```{r}
sdf <- read.transactions("http://bit.ly/SupermarketDatasetII")
```

## Lets preview our transaction dataset 

```{r}
head(sdf)
```
# Checking the dimensions of the data
```{r}
dim(sdf)

```
The dataset has 7501 transactions and 5729 columns


## Displaying the structure of our dataset
```{r}
str(sdf)
```

# Verifying the class of the object
```{r}
class(sdf)
```

# Previewing the first few items in the dataset
```{r}
inspect(sdf[1:6])
```

# Previewing the items in the dataset as if it were in a dataframe
```{r}
items<-as.data.frame(itemLabels(sdf))
colnames(items) <- "Item"
head(items, 5)
```



# Getting the summary statistics of the data
```{r}
summary(sdf)
```



# Exploring the frequency of some articles
```{r}
itemFrequency(sdf[, 7:11],type = "absolute")
```


# Producing a chart of frequencies and filtering to consider only items with a minimum percentage of support
```{r}
par(mfrow = c(1, 3))
```

# plot the frequency of items
```{r}
itemFrequencyPlot(sdf, topN = 10,col="grey")
itemFrequencyPlot(sdf, support = 0.1,col="orange")
```


OBSERVATION 

The top 10 most common items in the
transactions dataset are tea, wheat, mineral water, fat, yoghurt,  french fries, frozen vegetables


# Setting the parameters for our association analysis
```{r}
rules <- apriori (sdf, parameter = list(supp = 0.001, conf = 0.8))
```



Rules Observation
## We have a set of 271 rules 


## Observing rules built in our model

```{r}
inspect(rules[1:10])
```

## Lets building a apriori model with Min Support as 0.002 and confidence as 0.8.
```{r}
rules2 <- apriori (sdf,parameter = list(supp = 0.002, conf = 0.8))
```

OBSERVATION Rule 2
We have a set of 99 rules 

## Lets building apriori model with Min Support as 0.002 and confidence as 0.6.

```{r}
rules3 <- apriori (sdf, parameter = list(supp = 0.001, conf = 0.6))
```

OBSERVATION Rule 3, 
we have a set of 319 rules 


## We Perform an exploration of our model using the summary function
```{r}
summary(rules)
```


## Lets Observing rules built in our model i.e. first 7 model rules
```{r}
inspect(rules[1:7])
```
## We are Ordering the rules by a criteria such as the level of confidence then looking at the first five rules.
```{r}
rules<-sort(rules, by="confidence", decreasing=TRUE)
inspect(rules[1:5])
```







